---
title: "MY470 Computer Programming: The R Language"
author: "Main contributors: Siân Brooke, Friedrich Geiecke"
date: "Week 8 Lab: Example Answers"
output: html_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

### Loop Exercises

```python
# Exercise 1: Create a list that contains all integers from 1 to 100 (inclusive), 
# except that it has None for every integer that is divisible by 3 
# Your list should look like: [1, 2, None, 4, 5, None, 7, 8, None, 10, ...]

ls = []
for i in range(1, 101):
    if i % 3 == 0:
        ls.append(None)
    else:
        ls.append(i)
print(ls)
```

```{r}
# Answer:

ls <- rep(0)

for (x in 1:100) {
  if (x %% 3 == 0) {
    ls[x] <- NA
} else {
    ls[x] <- x } 
}
print(ls)


# With vectorization:

x <- 1:100
ls <- ifelse(x%%3==0, NA, x)
print(ls)
```

```python
# Exercise 2: Sum the even integers from the list below.

lst = [1, 3, 2, 4.5, 7, 8, 10, 3, 5, 4, 7, 3.33]

summ = 0
for i in lst:
    if i % 2 == 0:
        summ += i
        
print(summ)
```

```{r}
# Answer:

v1 <- c(1, 3, 2, 4.5, 7, 8, 10, 3, 5, 4, 7, 3.33)

summ <- 0
for (x in v1) {
  if (x %% 2 == 0) {
    summ <- summ + x
  }
}

print(summ)


# With vectorization:

summ <- sum(v1[v1 %% 2 == 0])
print(summ)
```


### Function Exercises

In Week 04, we looked at at defining our own functions. Translate the functions from the class exercises (below) into R.

```python
# Exercise 3

def zero_list(alist):
    """Takes a list and returns another list of the same length 
    that looks like [0, 0, 0, ...].
    """
    newlist = [0]*len(alist)  
    return newlist

mylist = [1, 2, 3]

zerolist = zero_list(mylist)  
print(mylist)
print(zerolist)
```

```{r}
# Answer:

# To read more about documenting functions in R see:
# https://cran.r-project.org/web/packages/roxygen2/vignettes/rd.html

#' Takes a vector and returns a new vector of the same length 
#' that looks like 0, 0, 0, ....
zero_vect <- function(a){
  newvector <- integer(length(a))
  return(newvector)
}

myvector <- c(1, 2, 3)

zerovector <- zero_vect(myvector)  
print(myvector) 
print(zerovector)
```

```python
# Exercise 4

scientists = {'Alan Turing': 'mathematician', 'Richard Feynman': 'physicist',
              'Marie Curie': 'chemist', 'Charles Darwin': 'biologist',
              'Ada Lovelace': 'mathematician', 'Werner Heisenberg': 'physicist'}

def print_professions(dic):
    """Takes a dictionary of {Name: profession} and prints
    'Name was a profession.'
    """
    for i in dic:
        print(i + ' was a ' + dic[i] + '.')
        
print_professions(scientists)
```

```{r}
# Answer:

scientists <- list('mathematician', 'physicist', 'chemist', 
                   'biologist', 'mathematician', 'physicist')

names(scientists) <- c('Alan Turing', 'Richard Feynman', 'Marie Curie', 
                       'Charles Darwin', 'Ada Lovelace', 'Werner Heisenberg')

print_professions <- function(science){
  for (x in names(science)) {
    print(paste0(x, ' was a ', science[[x]], "."))
  }}

print_professions(scientists)

# With vectorization
paste(names(scientists), 'was a', scientists)
```

### Vectorizing Exercises

Below we have the list comprehensions that we wrote in class in Week 3. For each Python snippet, create a chunk in R and translate the code into R.

```python
# Exercise 5: Create a new list containing the squares of the integers 
# in the list below.

lst = [1, 3, 2, 4.5, 7, 8, 10, 3, 5, 4, 7, 3.33]
ans = [i**2 for i in lst if type(i) == int] 
```

```{r}
# Answer: In R, vectors can only contain vectors of one data type.
# This means, to replicate the result, we instead look for
# whole numbers (i.e. divisible by 1)

lst <- c(1, 3, 2, 4.5, 7, 8, 10, 3, 5, 4, 7, 3.33)
ans <- lst[!lst%%1]^2
ans
```

```python
# Exercise 6: Using the function ifelse(), replicate the list comprehension below

import numpy as np

x = np.random.normal(0, 1, 100)
y = ["Positive" if i > 0 else "Negative" for i in x]
```

```{r}
# Answer:

x <- rnorm(100)

# ifelse(test, yes, no)
y <- ifelse(x < 0, "Negative", "Positive")
print(y)
```

```{r}
# Exercise 7: Vectorize the code below. This is a bad loop with 'growing' data.

# Set seed to compare the results
set.seed(42)

m <- 10
n <- 10

# Create matrix of normal random numbers
# replicate(num_of_replications, expression to evaluate)
# 10 x 10 normally distributed matrix
mymat <- replicate(m, rnorm(n))

# Transform into data frame
mydframe <- data.frame(mymat)

for (i in 1:m) {
  for (j in 1:n) {
    # For row, column, add some noise
    mydframe[i,j] <- mydframe[i,j] + 10*sin(0.75*pi)
  }
}
mydframe
```

```{r}
# Answer:

set.seed(42)
m <- 10
n<- 10

mymat <- replicate(m, rnorm(n)) 
mydframe <- data.frame(mymat)

# Very simple
mydframe <- mydframe + 10*sin(0.75*pi)
mydframe
``` 

Using the sample data specified below, use `?` with `apply()` to complete the steps below:

- Exercise 8: Create a column which is the mean of each row.
- Exercise 9: Create a row which is the sum of each column.
- Exercise 10: Create a function named `fun` that calculates the square of a number and convert the output to character if the character argument is set to `TRUE`.

```{r}

# Data frame
sample.df <- data.frame(x = 1:4, y = 5:8, z = 10:13)

# Answer to Exercise 8:
sample.df$row_mean <- apply(sample.df, MARGIN=1, mean)

# Answer to Exercise 9:
sample.df <- rbind(sample.df, apply(sample.df, MARGIN=2, sum))
print(sample.df)


# Answer to Exercise 10:
fun <- function(x, character=FALSE){
  sqx <- x^2
  if (character) {
    return(as.character(sqx))
  } else {
    return(sqx)
  }
}

# Check that it returns numeric
num = fun(3.4)
print(paste(num, class(num)))

# Check that it returns character
char = fun(3.4, character=TRUE)
print(paste(char, class(char)))

```
